Symmetric Circuits for Rank Gates

### 2018-01-30 sections/symmetric-circuits.tex
24th January 2018
30 Jan

As our aim is a circuit characterisation of $\FPR$, whose expressive power is
strictly greater than that of $\FPC$, we need to consider gates corresponding to
non-symmetric Boolean functions. In particular, we consider \emph{rank} gates
whose inputs are a matrix and whose result is invariant under row and column
permutations. To lead up to this, we first develop a general framework of
structured Boolean functions. These are functions whose inputs naturally encode
$\tau$-structures and where the output is invariant under the natural symmetries
of such structures. The\emph{matrix-invariant} functions are a natural
partiruclar case. We therefore define symmetric circuits in a general form where
gates can be labelled by \emph{isomorphism-invariant} structured functions.

@@

% AD - rewrote the paragraph to reduce the use of subscripts and superscripts.
Let $\tau = (R, S, \nu)$ be a many-sorted vocabulary. Let $D := \biguplus_{s \in
  S} D_{s} = \{(s,d) : d \in D_s\}$, be a disjoint union of non-empty sets,
indexed by $S$. For $R_i \in R$ let $R^D_i$ denote the trivial relation over the
many-sorted universe $D$, i.e.\ $R^D_i := D_{a_1} \times \ldots \times
D_{a_{r}}$, where $\nu(R_i) = (a_1, \ldots , a_{r)})$. There is an obvious
interpretation of $\tau$ over $D$ given by assigning each sort symbol in $\tau$
in the obvious manner and each relation symbol in $\tau$ to the trivial relation
over $D$. We call this \emph{the structure defined by $(\tau, D)$}, and denote
it by $\str(\tau, D)$.

The idea is to use the tuples in the relations of this structure as an index
set, and with the automorphisms of this structure defining a group action. Let
the \emph{index defined by the pair $(\tau, D)$} be the set $\biguplus_{R_i\in
  R} R^{D}_i$. We often use this set to index the input to a Boolean function,
as such we abuse notation and denote the index defined by $(\tau, D)$ by
$\ind(\tau, D)$, and call $\tau$ the vocabulary and $D$ the universe of
$\ind(\tau, D)$.

% GW: move sorted permutations into the background section
For the remainder of this subsection we fix a vocabulary $\tau$ and a
many-sorted set $D := \biguplus_{s \in S} D_{s}$, and let $X := \ind (\tau, D)$
and $G \leq \sym_{X}$. There is an obvious action of $G$ on $X$. There is also
an obvious action of the automorphism group $\str(\tau, D)$ on $X$ (with
permutations in $\str(\tau, D)$ acting element-wise on tuples in $X$). We denote
this group by $\aut(\tau, D)$, and identify it with the subgroup of $\sym_X$
that acts equivalently on $X$.

@@

We say that two functions $f: \ind(\tau, D) \rightarrow H$ and $g: \ind(\tau',
D') \rightarrow H$ are \emph{isomorphism-equivalent} if there exists an
isomorphism $\pi :\str(\tau, D) \rightarrow \str(\tau', D')$ such that $f (x) =
g (\pi (x))$ for all $x \in \str(\tau, D)$. It is easy to see that
isomorphism-equivalence is an equivalence relation.

We are often interested in the case where $X = \ind(\tau, D) = \ind(\tau', D')$.
We say that $f, g X \rightarrow H$ are \emph{$G$-equivalent} if there exists
$\sigma \in G$ such that $f = \sigma \cdot g$. We note that in this case $f$ and
$g$ are \emph{isomorphism-equivalent} if, and only if, they are $\aut(\tau,
G)$-equivalent.

In this paper we will often let the co-domain $H$ be the binary symbols
$\{0,1\}$, and so consider functions of the form $f : X \rightarrow \{0,1\}$. In
this case we may think of $f$ as defining for each relation $R^D_i$ a subset of
that relation, and thus determining a $\tau$-structure over the universe $D$. We
note that two functions $f, g: X \rightarrow \{0,1\}$ are isomorphism-equivalent
if, and only if, the two $\tau$-structures they determine are isomorphic.

@@

Structured functions give us a general framework for indexing the input sequence
to a Boolean function and associating with that function an appropriate symmetry
condition, or notion of equivalence. Let $F$ be a $(\tau, D)$-structured
function. We say $F$ is \emph{$G$-invariant} if it is constant on $G$-equivalent
input vectors, i.e. for all $G$-equivalent $\vec{x}, \vec{y} \in \{0,1\}^{X}$,
$F(\vec{x}) = F(\vec{y})$. We say $F$ is \emph{isomorphism-invariant} if it is
$\aut(\tau, D)$-invariant. Note that a structured function with vocabulary
$\tau$ is isomorphism-invariant if, and only if, it decides a property of
$\tau$-structures.

At the beginning of this subsection we defined three Boolean functions with
differently indexed inputs and corresponding symmetry conditions. We now look at
how these functions might be thought of as structured functions, and how the
symmetry conditions we discussed are naturally reproduced. Suppose $\tau$ is
single-sorted and consists of a single unary relation. Then any structured
function $F$ with vocabulary $\tau$ is isomorphism-invariant if, and only if, it
is symmetric. If instead the vocabulary of $F$ is the graph vocabulary, then $F$
is isomorphism-invariant if, and only if, $F$ decides a property of directed
graphs. Lastly, if the vocabulary of $F$ has just two sorts $(s_1, s_2)$ and one
binary relation with type $(s_1, s_2)$, then $F$ is isomorphism-invariant if,
and only if, it is matrix-invariant.

In each case a natural choice of the structure indexes the input to the
structured function appropriately and produces the appropriate symmetry
condition. It is worth mentioning that the graph-invariant case is not strictly
reproduced above, as the structured functions are rather defined over directed
graphs.

@@

In keeping with convention for Boolean functions, and for introducing simplicity
without a loss of generality, we assume that the universe of any structured
function is always a disjoint union of initial segments of the natural numbers,
i.e.\ if $F$ is structured function with universe $D_{s_1} \uplus \ldots \uplus
D_{s_p}$, then for all $s \in \{s_1 , \ldots , s_p\}$ there exists $d_s \in
\mathbb{N}$ such that $D_{s} = [d_s]$. We also assume that every $(\tau,
D)$-structured function has the property that every $a \in D$ appears in at
least one element of $\ind(\tau, D)$.

@@

We now define a \emph{$(\mathbb{B}, \tau)$-circuit} . This is a circuit that
computes a function on structures in the relational vocabulary $\tau$ and where
the gates are labelled by elements of the basis $\mathbb{B}$. Each element of
the basis is a $(\tau,D)$-structured function for some $\tau$ and $D$. Note
that, in the interests of keeping our definitions from getting excessively
complicated, we do not consider circuits computing functions over multi-sorted
vocabularies.

\begin{definition}[Circuits on Structures]
  Let $\mathbb{B}$ be a basis and $\tau$ be a relational vocabulary, we define a
  \emph{$(\mathbb{B}, \tau)$-circuit} $C_n$ computing a $q$-ary query $Q$ as a
  structure $\langle G, \Omega, \Sigma, \Lambda, L \rangle$.
  \begin{itemize}
    \setlength\itemsep{0mm}
  \item $G$ is called the set of gates of $C_n$ and $\vert C_n \vert := \vert G
    \vert$.
  \item $\Omega$ is an injective function from $[n]^q$ to $G$. The gates in the
    image of $\Omega$ are called the output gates. When $q = 0$, $\Omega$ is a
    constant function mapping to a single output gate.
  \item $\Sigma$ is a function from $G$ to $\mathbb{B} \uplus \tau \uplus
    \{0,1\} $ such that $\vert \Sigma^{-1} (0) \vert \leq 1$ and $\vert
    \Sigma^{-1} (1) \vert \leq 1$. Those gates mapped to $\tau \uplus \{0,1\}$
    are called input gates, with those mapped to $\tau$ called relational gates
    and those mapped $\{0,1\}$ called constant gates. Those gates mapped to
    $\mathbb{B}$ are called internal gates.
  \item $\Lambda$ is a sequence of injective functions $(\Lambda_{R_i})_{R_i \in
      R}$ such that $\Lambda_{R_i}$ maps each relational gate $g$ with $\Sigma
    (g) = R_i$ to the tuple $\Lambda_{R_i} (g) \in [n]^{r}$. When no ambiguity
    arises we write $\Lambda (g)$ for $\Lambda_{R_i} (g)$.
  \item $L$ associates with each internal gate $g$ a function $L(g):
    \ind(\Sigma(g)) \rightarrow G$ such that if we define a relation $W
    \subseteq G^{2}$ by $W(h_1,h_2)$ iff $h_2$ is an internal gate and $h_1$ is
    in the image of $L(h_2)$, then $(G, W)$ is a directed acyclic graph.
  \end{itemize}
\end{definition}

@@

Let $C_n = \langle G, \Omega, \Sigma, \Lambda, L \rangle$ be a $(\mathbb{B},
\rho)$-circuit. The $n$ is used to indicates that $C_n$ accepts structures of
size $n$. For a gate $g \in C_n$, we define the \emph{index} of $g$, denoted by
$\ind(g)$, as $\ind(\Sigma(g))$. We define the relation $W \subseteq G^2$ by
$W(h,g)$ if, and only if, $h \in \range(L(g))$. For each $g \in G$ we let
$W(\cdot, g) := \{h \in G : W(h,g)\}$ and $W(g, \cdot) := \{h \in G : W(g,h)\}$.
We call the elements of $W(\cdot, g)$ the \emph{children} of $g$ and the
elements of $W(g, \cdot)$ the \emph{parents} of $g$. We also abbreviate $W(g,
\cdot)$ by $H_g$.

We let the vocabulary and universe of $g$ be the vocabulary and universe of
$\Sigma(g)$, respectively, and denote the vocabulary of $g$ by $\vocab{g}$ and
the universe by $\universe{g}$. We also let $\aut(g) := \aut(\tau, D)$ and
$\str(g) := \str(\tau, D)$.

@@

Let $\tau$ be a relational vocabulary, $\mathcal{A}$ a $\tau$-structure over an
$n$-element universe $U$ and $\gamma \in [n]^{\underline{U}}$. Let $\gamma
\mathcal{A}$ be the structure on universe $[n]$ formed by mapping $U$ in
accordance with $\gamma$. The evaluation of a $(\mathbb{B}, \tau)$-circuit $C_n$
computing a $q$-ary query $Q$ proceeds by recursively evaluating the gates in
the circuit. The evaluation of the gate $g$ for the bijection $\gamma$ and input
$\mathcal{A}$ is denoted by $C[\gamma \mathcal{A}](g)$, and is given as follows
\begin{enumerate}
  \setlength\itemsep{0mm}
\item If $g$ is a constant gate then it evaluates to the bit given by
  $\Sigma(g)$.
\item If $g$ is a relational gate then $g$ evaluates to true iff $\gamma
  \mathcal{A} \models \Sigma(g)(\Lambda (g))$.
\item If $g$ is an internal gate such that $\Sigma (g)$ let $L^{\gamma}_g:
  \ind(g) \rightarrow \{0,1\}$ be defined by $L^{\gamma}_g(x) = C[\gamma
  \mathcal{A}](L(g)(x))$, for all $x \in \ind(g)$. Then $g$ evaluates to true
  iff $\Sigma(g) (L^{\gamma}_g) = 1$.
\end{enumerate}
We say that $C_n$ defines the $q$-ary query $Q \subseteq U^q$ under $\gamma$
where $\vec{a} \in Q$ if, and only if, $C_n[\gamma \mathcal{A}](\Omega (\gamma
\vec{a})) = 1$.

@@

\begin{definition}[Invariant Circuit]
  Let $C_n$ be a $(\mathbb{B}, \tau)$-circuit, computing some $q$-ary query. We
  say $C_n$ is \emph{invariant} if for every $\tau$-structure $\mathcal{A}$ of
  size $n$, $\vec{a} \in U^{q}$, and $\gamma_1, \gamma_2: [n]^{\underline{U}}$
  we have that $C[\gamma_1 \mathcal{A}](\Omega (\gamma_1 \vec{a})) = C[\gamma_2
  \mathcal{A}](\Omega (\gamma_2 \vec{a}))$.
\end{definition}

In other words $C_n$ is invariant if the evaluation of the circuit on any
$n$-element structure $\mathcal{A}$ does not depend on the particular map
$\gamma$ that takes the universe of the structure to $[n]$. In general, the
evaluation of a circuit depends on the presentation of the structure (i.e.\ the
bijection $\gamma$). If a family of $(\mathcal{B}, \tau)$-circuits $\mathcal{C}$
is invariant it follows that the query computed is a $q$-query on
$\tau$-structures. Thus if $q = 0$ then $\mathcal{C}$ computes a property of
$\tau$-structures. The following lemma allows us to recast this notion in terms
of the language developed in this paper.

\begin{lem}
  Let $C_n$ be a $(\mathbb{B}, \tau)$-circuit. The function computed by $C_n$ is
  $\tau$-invariant if, and only if, $C_n$ is an invariant circuit.
\end{lem}
% AD - does this require a few lines of proof? GW - To be added. I have not done
% this yet.

% \begin{definition}[Automorphism]
%   let $C = \langle G, \Omega, \Sigma, \Lambda, L\rangle$ be a
%   $(\mathbb{B},\tau)$-circuit computing a $q$-ary query on structures of size
%   $n$, and where $\mathbb{B}$ is a basis of isomorphism-invariant structured
%   functions. Let $\sigma \in \sym_n$ and $\pi: G \rightarrow G$ be a bijection
%   such that
%   \begin{itemize}
%     \setlength\itemsep{0mm}
%   \item for all output tuples $x \in [n]^q$, $\pi \Omega (x) = \Omega (\sigma
%     x)$,
%   \item for all gates $g \in G$, let $\Sigma (g) = \Sigma (\pi g)$,
%   \item for each relational gate $g \in G$, $\sigma \Lambda (g) = \Lambda (\pi
%     g)$, and
%   \item For each pair of gates $g, h \in G$ $W(h,g)$ iff $W(\pi h, \pi g)$ and
%     for each internal gate $g$ we have that $L(\pi g)$ and $ \pi \cdot L(g)$
%     are isomorphism-equivalent.
%   \end{itemize}
%   We call $\pi$ an \emph{automorphism} of $C$, and we say that $\sigma$
%   \emph{extends to an automorphism} $\pi$. The group of automorphisms of $C$
%   is called $\aut (C)$.
% \end{definition}

>>>>

As our aim is a circuit characterisation of $\FPR$, whose expressive power is
strictly greater than that of $\FPC$, we need to consider gates corresponding to
non-symmetric Boolean functions. In particular, we consider \emph{rank} gates
whose inputs are a matrix and whose result is invariant under row and column
permutations. To lead up to this, we first develop a general framework of
structured Boolean functions. These are functions whose inputs naturally encode
$\rho$-structures and where the output is invariant under the natural symmetries
of such structures. The\emph{matrix-invariant} functions are a natural
partiruclar case. We therefore define symmetric circuits in a general form where
gates can be labelled by \emph{isomorphism-invariant} structured functions.

@@

% AD - rewrote the paragraph to reduce the use of subscripts and superscripts.
Let $\tau := (R, S, \nu)$ be a many-sorted vocabulary. Let $D := \biguplus_{s \in
  S} D_{s} = \{(s,d) : d \in D_s\}$, be a disjoint union of non-empty sets,
indexed by $S$. For $R_i \in R$ let $R^D_i$ denote the trivial relation over the
many-sorted universe $D$, i.e.\ $R^D_i := D_{a_1} \times \ldots \times
D_{a_{r}}$, where $\nu(R_i) = (a_1, \ldots , a_{r)})$. There is an obvious
interpretation of $\tau$ over $D$ given by assigning each sort symbol in $\tau$
in the obvious manner and each relation symbol in $\tau$ to the trivial relation
over $D$. We call this \emph{the structure defined by $(\tau, D)$}, and denote
it by $\str{\tau, D}$.

The idea is to use the tuples in the relations of this structure as an index
set, and with the automorphisms of this structure defining a group action. Let
the \emph{index defined by the pair $(\tau, D)$} be the set $\biguplus_{R_i\in
  R} R^{D}_i := \{ (\vec{a}, R) : \vec{a} \in R^{D}, \, R \in \tau \} $. We
often use this set to index the input to a Boolean function, and as such we
denote the index defined by $(\tau, D)$ by $\ind(\tau, D)$, and call $\tau$ the
vocabulary and $D$ the universe of $\ind(\tau, D)$.

We abbreviate $(\vec{a}, R) \in \ind(\tau, D)$ by $\vec{a}_R$, and refer
to $R$ as \emph{the type} of $\vec{a}_R$. We often abuse notation and treat
$\vec{a}_R$ as the tuple $\vec{a}$ in $R^D$. In the event that there is only one
relational symbol in $\tau$ we identify $\vec{a}_R$ with $\vec{a}$.

% Although we often use the convention of identifying the elements of the
% disjoint union with the elements of the sets from which the union was formed
% (i.e. identify the elements of $\ind(\tau, D)$ with the tuples that appear in
% a some relation in $\str(\tau, D)$), it is important to note that $\ind(\tau,
% D)$ is formally a disjoint union, and as such each tuple in this set is
% associated with an identifiable relation symbol. We refer to $R$ as the type

% GW: move sorted permutations into the background section
For the remainder of this subsection we fix a vocabulary $\tau$ and a
many-sorted set $D := \biguplus_{s \in S} D_{s}$, let $X := \ind (\tau, D)$ and $G \leq \sym_{X}$.
There is an obvious action of $G$ on $X$. There is also an action of the
automorphism group $\str{\tau, D}$ on $X$ defined for the automorphism $\sigma$
by $\sigma \cdot \vec{a}_R := (\sigma \vec{a})_R$. We denote the group of
automorphisms of $\str{\tau, D}$ by $\aut(\tau, D)$, and identify it with the
subgroup of $\sym_X$ that acts equivalently on $X$.

@@

We say that two functions $f: \ind(\tau, D) \rightarrow H$ and $g: \ind(\tau,
D') \rightarrow H$ are \emph{isomorphism-equivalent} if there exists an
isomorphism $\pi :\str{\tau, D} \rightarrow \str{\tau, D'}$ such that $f (\vec{a}_R) =
g (\pi \vec{a}_R)$ for all $\vec{a}_R \in \ind(\tau, D)$. It is easy to see that
isomorphism-equivalence is an equivalence relation.

We are often interested in the case where $X = \ind(\tau, D) = \ind(\tau, D')$.
We say that $f, g: X \rightarrow H$ are \emph{$G$-equivalent} if there exists
$\sigma \in G$ such that $f = \sigma \cdot g$. We note that in this case $f$ and
$g$ are isomorphism-equivalent if, and only if, they are $\aut(\tau,
D)$-equivalent.

In this paper we will often take the co-domain $H$ to be $\{0,1\}$, and so $f :
X \rightarrow \{0,1\}$. In this case we may think of $f$ as defining for each
relation $R^D_i$ a subset of that relation, and so determining a
$\tau$-structure over the universe $D$. We note that two functions $f, g: X
\rightarrow \{0,1\}$ are isomorphism-equivalent if, and only if, the two
$\tau$-structures they determine are isomorphic.

@@

Structured functions give us a general framework for indexing the input sequence
to a Boolean function and associating with that function an appropriate symmetry
condition, or notion of equivalence. Let $F$ be a $(\tau, D)$-structured
function. We say $F$ is \emph{$G$-invariant} if it is constant on $G$-equivalent
input vectors, i.e. for all $G$-equivalent $f, g \in \{0,1\}^{X}$, $F(f) =
F(g)$. We say $F$ is \emph{isomorphism-invariant} if it is $\aut(\tau,
D)$-invariant. Note that a structured function with vocabulary $\tau$ is
isomorphism-invariant if, and only if, it decides a property of
$\tau$-structures.

At the beginning of this subsection we defined three Boolean functions with
differently indexed inputs and corresponding symmetry conditions. We now look at
how these functions might be thought of as structured functions, and how the
symmetry conditions we discussed are naturally reproduced. Suppose $\tau$ is
single-sorted and consists of a single unary relation symbol. Then any structured
function $F$ with vocabulary $\tau$ is isomorphism-invariant if, and only if, it
is symmetric. If instead the vocabulary of $F$ is the graph vocabulary, then $F$
is isomorphism-invariant if, and only if, $F$ decides a property of directed
graphs. Lastly, if the vocabulary of $F$ has just two sorts $(s_1, s_2)$ and one
binary relation with type $(s_1, s_2)$, then $F$ is isomorphism-invariant if,
and only if, it is matrix-invariant.

In each case a natural choice of $\tau$ (and $D$) indexes the input to the
structured function appropriately and produces the appropriate symmetry
condition. It is worth mentioning that the graph-invariant case is not strictly
reproduced using structured functions. This is because structured functions
index the input string using the elements of the relations of the complete
$\tau$-structure (in this case, edges in the complete graph). As such, the input
string is indexed by ordered pairs of elements, rather than two-element
subsets, and as such the symmetry condition on the relation is not enforced. As
such, in the structured function case the function decides properties of
directed graphs, rather than ordinary graphs. We should note, however, that the symmetry
of the edge relation is a property of directed-graph, and as such can be decided
by an isomorphism-invariant structured function over the graph vocabulary.

% , and any
% input sequence encodes a sub-relation of that complete relation. As such, the
% function  This is because in the event that $\tau := {E}$ (where $E$ is
% a binary relation, thought of as the edge relation) there is no assurance that
% the function $f : \ind
% (\tau, D) \rightarrow \{0,1\}$ is encoding

% as the structured functions are rather defined over directed
% graphs.

@@

In keeping with convention for Boolean functions, and for introducing simplicity
without a loss of generality, we assume that the universe of any structured
function is always a disjoint union of initial segments of the natural numbers,
i.e.\ if $F$ is structured function with universe $D_{s_1} \uplus \ldots \uplus
D_{s_p}$, then for all $s \in \{s_1 , \ldots , s_p\}$ there exists $d_s \in
\nats{N}$ such that $D_{s} = [d_s]$. We also assume that every $(\tau,
D)$-structured function has the property that every $a \in D$ appears in at
least one element of $\ind(\tau, D)$.

@@

We now define a \emph{$(\mathbb{B}, \rho)$-circuit} . This is a circuit that
computes a function on structures in the relational vocabulary $\rho$ and where
the gates are labelled by elements of the basis $\mathbb{B}$ consisting of
structured functions. Note that, in the interests of keeping our definitions
from getting excessively complicated, we do not consider circuits computing
functions over multi-sorted vocabularies.

\begin{definition}[Circuits on Structures]
  Let $\mathbb{B}$ be a basis and $\rho$ be a relational vocabulary, we define a
  \emph{$(\mathbb{B}, \rho)$-circuit} $C$ of order $n$ computing a $q$-ary query
  $Q$ as a structure $\langle G, \Omega, \Sigma, \Lambda, L \rangle$.
  \begin{itemize}
    \setlength\itemsep{0mm}
  \item $G$ is called the set of gates of $C$ and $\vert C \vert := \vert G
    \vert$.
  \item $\Omega$ is an injective function from $[n]^q$ to $G$. The gates in the
    image of $\Omega$ are called the output gates. When $q = 0$, $\Omega$ is a
    constant function mapping to a single output gate.
  \item $\Sigma$ is a function from $G$ to $\mathbb{B} \uplus \rho \uplus
    \{0,1\} $ such that $\vert \Sigma^{-1} (0) \vert \leq 1$ and $\vert
    \Sigma^{-1} (1) \vert \leq 1$. Those gates mapped to $\rho \uplus \{0,1\}$
    are called input gates, with those mapped to $\rho$ called relational gates
    and those mapped $\{0,1\}$ called constant gates. Those gates mapped to
    $\mathbb{B}$ are called internal gates.
  \item $\Lambda$ is a sequence of injective functions $(\Lambda_{R_i})_{R_i \in
      R}$ such that $\Lambda_{R_i}$ maps each relational gate $g$ with $\Sigma
    (g) = R_i$ to the tuple $\Lambda_{R_i} (g) \in [n]^{r}$. When no ambiguity
    arises we write $\Lambda (g)$ for $\Lambda_{R_i} (g)$.
  \item $L$ associates with each internal gate $g$ a function $L(g):
    \ind(\Sigma(g)) \rightarrow G$ such that if we define a relation $W
    \subseteq G^{2}$ by $W(h_1,h_2)$ iff $h_2$ is an internal gate and $h_1$ is
    in the image of $L(h_2)$, then $(G, W)$ is a directed acyclic graph.
  \end{itemize}
\end{definition}

@@

Let $C = \langle G, \Omega, \Sigma, \Lambda, L \rangle$ be a $(\mathbb{B},
\rho)$-circuit of order $n$. The order of a circuit denotes the fact that it
takes as input structures of size $n$. We often consider families of circuits of
the form $(C_n)_{n \in \nats}$, and here we assume that each $C_n$ is a circuit
of order $n$. For a gate $g \in C$, we define the \emph{index} of $g$, denoted
by $\ind(g)$, as $\ind(\Sigma(g))$. We define the relation $W \subseteq G^2$ by
$W(h,g)$ if, and only if, $h \in \range(L(g))$. For each $g \in G$ we let
$W(\cdot, g) := \{h \in G : W(h,g)\}$ and $W(g, \cdot) := \{h \in G : W(g,h)\}$.
We call the elements of $W(\cdot, g)$ the \emph{children} of $g$ and the
elements of $W(g, \cdot)$ the \emph{parents} of $g$. We also abbreviate $W(g,
\cdot)$ by $H_g$.

We let the vocabulary and universe of $g$ be the vocabulary and universe of
$\Sigma(g)$, respectively, and denote the vocabulary of $g$ by $\vocab{g}$ and
the universe by $\universe{g}$. We also let $\aut(g) := \aut(\vocab{g}, \universe{g})$ and
$\str{g} := \str{\vocab{g}, \universe{g}}$.

@@

Let $\rho$ be a relational vocabulary, $\mathcal{A}$ a $\rho$-structure over an
$n$-element universe $U$ and $\gamma \in [n]^{\underline{U}}$. Let $\gamma
\mathcal{A}$ be the structure on universe $[n]$ formed by mapping $U$ in
accordance with $\gamma$. The evaluation of a $(\mathbb{B}, \rho)$-circuit $C$
of order $n$ computing a $q$-ary query $Q$ proceeds by recursively evaluating
the gates in the circuit. The evaluation of the gate $g$ for the bijection
$\gamma$ and input $\mathcal{A}$ is denoted by $C[\gamma \mathcal{A}](g)$, and
is given as follows
\begin{enumerate}
  \setlength\itemsep{0mm}
\item If $g$ is a constant gate then it evaluates to the bit given by
  $\Sigma(g)$,
\item if $g$ is a relational gate then $g$ evaluates to true iff $\gamma
  \mathcal{A} \models \Sigma(g)(\Lambda (g))$, and
\item if $g$ is an internal gate such that $\Sigma (g)$ let $L^{\gamma}_g:
  \ind(g) \rightarrow \{0,1\}$ be defined by $L^{\gamma}_g(x) = C[\gamma
  \mathcal{A}](L(g)(x))$, for all $x \in \ind(g)$. Then $g$ evaluates to true
  if, and only if, $\Sigma(g) (L^{\gamma}_g) = 1$.
\end{enumerate}
We say that $C$ defines the $q$-ary query $Q \subseteq U^q$ under $\gamma$ where
$\vec{a} \in Q$ if, and only if, $C[\gamma \mathcal{A}](\Omega (\gamma \vec{a}))
= 1$.

@@

\begin{definition}[Invariant Circuit]
  Let $C$ be a $(\mathbb{B}, \rho)$-circuit of order $n$, computing some $q$-ary
  query. We say $C$ is \emph{invariant} if for every $\rho$-structure
  $\mathcal{A}$ of size $n$, $\vec{a} \in U^{q}$, and $\gamma_1, \gamma_2:
  [n]^{\underline{U}}$ we have that $C[\gamma_1 \mathcal{A}](\Omega (\gamma_1
  \vec{a})) = C[\gamma_2 \mathcal{A}](\Omega (\gamma_2 \vec{a}))$.
\end{definition}

It follows that the circuit $C$ of order $n$ is invariant if, and only if, the
evaluation of the circuit for any input $n$-element structure $\mathcal{A}$ does
not depend on the bijection $\gamma$. In general, the evaluation of a circuit
depends on the presentation of the structure (i.e.\ the bijection $\gamma$). If
a family of $(\mathcal{B}, \rho)$-circuits $\mathcal{C}$ is invariant it follows
that the query computed is a $q$-query on $\rho$-structures. Thus if $q = 0$
then $\mathcal{C}$ computes a property of $\rho$-structures. The following lemma
allows us to recast this notion in terms of the language developed in this
paper.

\begin{lem}
  Let $C$ be a $(\mathbb{B}, \rho)$-circuit of order $n$ computing a $0$-ary
  query. The function computed by $C$ is isomorphism-invariant if, and only if,
  $C$ is an invariant circuit.
\end{lem}
\begin{proof}
  We first note that, since $C$ computes a $0$-ary query, there is exactly one
  output gate. We call this gate $g_o$. We can associate with each relational
  gate $g$ in $C$ a unique pair $(\vec{a}, R)$, where $R := \Sigma (g)$ is a
  relational symbol in $\tau$ and $\vec{a} := \Lambda_R(g)$ is an an element of
  $R^{[n]}$. Thus we can think of the function $F_C$ computed by the circuit as
  having it's input string indexed by the elements of $\ind(\rho, [n])$, and so
  think of $F_C$ as a $(\rho, [n])$-structured function.

  % The value of $F_C (f)$, for some element $f \in
  % \{0,1\}^{\ind(\rho, [n])}$, is computed by assigning to each relational gate
  % $(\vec{a}, R)$ the value $f(\vec{a}, R)$, recursively evaluating the
  % circuit,
  % and letting $F_C (f)$ be the result of evaluating $g_{\text{out}}$.

  We now present a number of observations, and then combine these observations
  to prove both directions of the hypothesis. Let $\mathcal{B}$ be a
  $\rho$-structure of size $n$ over the universe $[n]$. We can define a function
  $f_{\mathcal{B}} : \ind(\rho, [n]) \rightarrow {0,1}$ such that
  $f_{\mathcal{B}} (\vec{a}_R) = 1$ if, and only if, $\vec{a} \in
  R^{\mathcal{B}}$. Let $f : \ind(\rho, [n]) \rightarrow \{0,1\}$. We can define
  the $\rho$-structure $\mathcal{B}_f$ over the universe $[n]$ such that for all
  $R \in \rho$, we have $\vec{a} \in R^{\mathcal{B}_f}$ if, and only if,
  $f(\vec{a}_R) = 1$. It is easy to see that the functions $\mathcal{B} \mapsto
  f_{\mathcal{B}}$ and $f \mapsto \mathcal{B}_f$ are inverse to one another, and
  so define a bijection. Moreover, we notice that for all $\sigma \in \sym_n$ we
  have that $f_{\sigma \mathcal{B}}(\vec{a}_R) = 1$ if, and only if, $\vec{a}
  \in R^{\sigma \mathcal{B}}$ if, and only if, $\sigma^{-1} \vec{a} \in
  R^{\mathcal{B}}$ if, and only if, $\sigma ^{-1} f_{\mathcal{B}}(\vec{a}_R) =
  f_{\mathcal{B}}(\sigma^{-1} \vec{a}_R)= 1$.

  Let $\mathcal{A}$ be a $\rho$-structure of size $n$ over the universe $U$, and
  let $\gamma_1, \gamma_2 \in [n]^{\underline{U}}$. From the definition of $F_C$
  it follows that for $\gamma \in \sym_n$ we have that $F_C (f_{\gamma
    \mathcal{A}})$ if, and only if, $C[\gamma \mathcal{A}](g_{\text{o}}) = 1$.

  Let $\gamma_1, \gamma_2 \in [n]^{\underline{U}}$ and let $\sigma \in \sym_n$
  be such that $\gamma_1 = \sigma \gamma_2$. We have that
  \begin{align*}
    C[\gamma_1 \mathcal{A}](g_{\text{o}}) = C[\gamma_2 \mathcal{A}](g_{\text{o}}) & \Leftrightarrow  F_C(f_{\gamma_1 \mathcal{A}}) = F_C (f_{\gamma_2 \mathcal{A}}) \\
                                                                                  & \Leftrightarrow F_C(f_{\gamma_1 \mathcal{A}}) = F_C (f_{\sigma^{-1}\gamma_1 \mathcal{A}}) \\
                                                                                  & \Leftrightarrow F_C(f_{\gamma_1 \mathcal{A}}) = F_C (\sigma f_{\gamma_1 \mathcal{A}}).
  \end{align*}

  We now combine the above observations to prove the result. Suppose $F_C$ is
  isomorphism-invariant. Let $\gamma_1, \gamma_2 \in [n]^{\underline{U}}$ and
  let $\sigma \in \sym_n$ be such that $\gamma_1 = \sigma \gamma_2$. Then, from
  isomorphism-invariance, we have $F_C(f_{\gamma_1 \mathcal{A}}) = F_C (\sigma
  f_{\gamma_1 \mathcal{A}})$, and so $C[\gamma_1 \mathcal{A}](g_{\text{o}}) =
  C[\gamma_2 \mathcal{A}](g_{\text{o}})$.

  Suppose $C$ is invariant. Fix the bijection $\gamma_1 \in
  [n]^{\underline{U}}$. Let $\sigma \in \sym_n$ and $f: \ind(\rho, [n])
  \rightarrow \{0,1\}$. Let $\mathcal{A} = \gamma^{-1}_1 \mathcal{B}_f$ and
  $\gamma_2 := \sigma^{-1}\gamma_1 $. Since $C$ is invariant we have $C[\gamma_1
  \mathcal{A}](g_{\text{o}}) = C[\gamma_2\mathcal{A}](g_{\text{o}})$, and so
  $F_C (\sigma f_{\gamma_1 \mathcal{A}}) = F_C(f_{\gamma_1 \mathcal{A}})$. But
  $f_{\gamma_1 \mathcal{A}} = f_{\mathcal{B}_f} = f$, and so $F_C (\sigma f) =
  F_C (f)$. It follows that $F_C$ is isomorphism-invariant.

  % Let $f_{\gamma \mathcal{A}} : \ind(\rho, [n]) \rightarrow {0,1}$ be defined
  % by
  % $f (\vec{a}_R) = 1$ if, and only if, $\vec{a} \in R^{\gamma \mathcal{A}}$.
  % So
  % $C[\gamma_1 \mathcal{A}](g_{\text{out}}) = F_C (f_{\gamma_1 \mathcal{A}}) =
  % F_C(f_{\sigma \gamma_2 \mathcal{A}}) = F_C (\sigma f_{\gamma_2 \mathcal{A}})
  % =
  % F_C (f_{\gamma_2 \mathcal{A}}) = C[\gamma_2 \mathcal{A}](g_{\text{out}})$.

  % are labelled by relational symbols from $\tau$ and elements of the
  % complete-re

  % Let $X := \ind(\rho, [n])$. We can define an injection $f_R$ that maps the
  % relational gate $g$ to $\vec{a}_R \in X$, where $R:= \Sigma(g)$ and $\vec{a}
  % := \Lamda_R(g)$. function computed by $C$ $F_C : \{0,1\}^{X} \rightarrow
  % \{0,1\}$.
\end{proof}
% AD - does this require a few lines of proof? GW - To be added. I have not done
% this yet.
